---
title: "Benchmark Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmark Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 8,
    fig.height = 5
)
```

This vignette compares ten context-retrieval strategies implemented in
**rrlmgraph-bench** on a suite of realistic R-project tasks.  Results
were pre-computed by the `run-benchmark` GitHub Actions workflow in
[rrlmgraph-bench](https://github.com/davidrsch/rrlmgraph-bench) and
summarised here — no benchmark re-run is needed to render this page.

## Strategies

| Strategy | Description |
|----------|-------------|
| `graph_rag_tfidf` | Graph-RAG: BFS from TF-IDF seed nodes |
| `graph_rag_tfidf_noseed` | Graph-RAG: TF-IDF traversal seeded by the query itself |
| `graph_rag_ollama` | Graph-RAG: BFS from Ollama vector-similarity seed nodes |
| `graph_rag_mcp` | Graph-RAG: traversal via the rrlmgraph-mcp MCP server |
| `graph_rag_agentic` | RLM-Graph: LLM writes R code in sandboxed REPL (Algorithm 1, arXiv:2512.24601); only compact stdout returned per iteration |
| `full_files` | Entire source files (**upper baseline**) |
| `bm25_retrieval` | BM25 keyword retrieval (no graph) |
| `term_overlap` | Word-overlap ranking (no graph) |
| `no_context` | No context — LLM answers from training data (**lower baseline**) |
| `random_k` | Five randomly sampled code chunks (random baseline) |

## Results

```{r load-summary}
# When a vignette is built, cwd = vignettes/ so this relative path works
# both during R CMD check and pkgdown site builds.
precomputed_path <- file.path("precomputed", "benchmark_summary.rds")
if (file.exists(precomputed_path)) {
    bench <- readRDS(precomputed_path)
} else {
    bench <- NULL
}
```

```{r show-summary, results = "asis"}
if (is.null(bench)) {
    cat(
        "> **Note:** Benchmark results not yet available.\n",
        "> Trigger the `run-benchmark` workflow in",
        "[rrlmgraph-bench](https://github.com/davidrsch/rrlmgraph-bench/actions)",
        "to generate data, then run the `precompute-vignettes` workflow here.\n"
    )
}
```

```{r summary-table, eval = !is.null(bench)}
stats <- bench$stats
knitr::kable(
    stats$summary[, c(
        "strategy", "mean_score", "sd_score", "ci_lo_95", "ci_hi_95",
        "mean_total_tokens", "hallucination_rate", "n"
    )],
    digits = 3,
    caption = paste0(
        "Mean score ± 95 % CI per strategy.  ",
        "Captured: ", bench$meta$generated_at, ".  ",
        "rrlmgraph v", bench$meta$rrlmgraph_version, " / ",
        "rrlmgraphbench v", bench$meta$rrlmgraphbench_version, "."
    ),
    col.names = c(
        "Strategy", "Mean score", "SD", "CI lower", "CI upper",
        "Mean tokens", "Hallucination rate", "N"
    )
)
```

```{r score-plot, eval = !is.null(bench), fig.cap = "Mean score ± 95 % CI by strategy.  Dashed line = full_files baseline."}
stats <- bench$stats
df <- stats$summary
df$strategy <- factor(df$strategy, levels = df$strategy[order(df$mean_score)])
baseline <- df$mean_score[df$strategy == "full_files"]

op <- par(mar = c(5, 10, 3, 2))
plot(
    df$mean_score,
    seq_len(nrow(df)),
    xlim = c(0, 1),
    xlab = "Mean score",
    ylab = "",
    yaxt = "n",
    pch = 16,
    col = ifelse(df$strategy == "graph_rag_tfidf", "steelblue",
        ifelse(df$strategy == "graph_rag_mcp", "seagreen", "grey40")
    ),
    main = "Strategy comparison"
)
axis(2, at = seq_len(nrow(df)), labels = df$strategy, las = 1, cex.axis = 0.85)
abline(v = baseline, lty = 2, col = "grey60")
arrows(
    df$ci_lo_95, seq_len(nrow(df)),
    df$ci_hi_95, seq_len(nrow(df)),
    length = 0.05, angle = 90, code = 3,
    col = ifelse(df$strategy == "graph_rag_tfidf", "steelblue",
        ifelse(df$strategy == "graph_rag_mcp", "seagreen", "grey40")
    )
)
par(op)
```

## Token efficiency

The Token Efficiency Ratio (TER) measures score gain per token relative
to the `full_files` baseline.  Values > 1 indicate better score per
token used.

```{r ter-table, eval = !is.null(bench)}
ter_df <- data.frame(
    strategy = names(bench$stats$ter),
    TER = round(bench$stats$ter, 3),
    stringsAsFactors = FALSE
)
knitr::kable(ter_df, row.names = FALSE, caption = "Token Efficiency Ratio (TER)")
```

## Pairwise tests

```{r pairwise, eval = !is.null(bench)}
pw <- bench$stats$pairwise
pw$significant <- pw$p_bonferroni < 0.05
knitr::kable(
    pw[, c("strategy_1", "strategy_2", "p_bonferroni", "cohens_d", "significant")],
    digits    = 4,
    col.names = c("Strategy A", "Strategy B", "p (Bonferroni)", "Cohen's d", "Sig."),
    caption   = "Pairwise Welch t-tests (Bonferroni-corrected)."
)
```

## Methodology

Full task definitions and scoring rubrics are in the
[rrlmgraph-bench](https://github.com/davidrsch/rrlmgraph-bench)
repository.  Each task is scored 0–1 using `compute_benchmark_statistics()`
from the **rrlmgraphbench** package.  Hallucinations are detected by
`count_hallucinations()`.
