---
title: "MCP Server Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MCP Server Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The **rrlmgraph-mcp** companion package exposes the project knowledge
graph via the [Model Context Protocol](https://modelcontextprotocol.io/)
(MCP), enabling any MCP-capable editor or AI agent to query graph context
without running R code directly.

## Architecture

```
┌─────────────────┐        JSON-RPC / stdio        ┌─────────────────┐
│   VS Code /     │ ◄────────────────────────────► │  rrlmgraph-mcp  │
│  GitHub Copilot │                                 │  (Node.js MCP   │
│   (MCP client)  │                                 │    server)      │
└─────────────────┘                                 └────────┬────────┘
                                                             │  SQLite
                                                    ┌────────▼────────┐
                                                    │  .rrlmgraph/    │
                                                    │   graph.db      │
                                                    └─────────────────┘
```

The SQLite database is written by `export_to_sqlite()` in **rrlmgraph**
and read at query time by the MCP server — no R process needs to be
running during IDE sessions.

## Installation

Install the MCP server globally via npm:

```{r install-mcp}
# In a terminal:
# npm install -g rrlmgraph-mcp
```

Or use `npx` for a one-off run:

```{r npx}
# npx rrlmgraph-mcp --db path/to/.rrlmgraph/graph.db
```

## Exporting the graph to SQLite

```{r export}
library(rrlmgraph)

graph <- build_rrlm_graph("path/to/your/project")
export_to_sqlite(graph) # writes .rrlmgraph/graph.db
```

## Configuring VS Code

Add the following to `.vscode/mcp.json` (or your user-level
`settings.json` under `github.copilot.chat.mcp.servers`):

```json
{
  "servers": {
    "rrlmgraph": {
      "type": "stdio",
      "command": "npx",
      "args": [
        "rrlmgraph-mcp",
        "--db",
        "${workspaceFolder}/.rrlmgraph/graph.db"
      ]
    }
  }
}
```

After saving, GitHub Copilot Chat will list **rrlmgraph** as an available
MCP tool.  Ask it questions like:

> *"Which functions call `prepare_data()`?"*
> *"What context is relevant for the data validation pipeline?"*

## Available MCP tools

| Tool | Description |
|------|-------------|
| `query_context` | Relevance-guided BFS context window for a coding task |
| `get_node_info` | Full metadata, callers, callees, and source for a named node |
| `list_functions` | Functions ranked by PageRank, with optional file-path filter |
| `find_callers` | All functions that call a given function (incoming CALLS edges) |
| `find_callees` | All functions called by a given function (outgoing CALLS edges) |
| `search_nodes` | Full-text keyword search across node names, bodies, and docs |
| `add_task_trace` | Record accepted / rejected task outcome to improve future retrieval |
| `rebuild_graph` | Trigger an Rscript rebuild of the project graph in-process |

## Keeping the database fresh

Re-run `export_to_sqlite()` after editing R source files, or use
`update_graph_incremental()` followed by `export_to_sqlite()` for large
projects:

```{r update}
graph <- update_graph_incremental(
  graph,
  changed_files = "R/my_function.R"
)
export_to_sqlite(graph)
```

A `{targets}` pipeline or a Git pre-commit hook that runs this
two-step command keeps the graph in sync with every commit.

## Further reading

- [rrlmgraph-mcp source](https://github.com/davidrsch/rrlmgraph-mcp)
- [Model Context Protocol specification](https://modelcontextprotocol.io/specification)
- [VS Code MCP documentation](https://code.visualstudio.com/docs/copilot/chat/mcp-servers)
